### 2022-10-06 AISTATS2023/paper.tex
added analysis to the real example

The California Housing is a largely-studied dataset with approximately
\(20000\) training instances, making it appropriate for robust
Monte-Carlo approximations. The dataset contains \(D=8\) features
describing building blocks of California, like latitude, longitude,
total number of people residing in the block etc., and the target is
the median house value inside the block. We exclude instances with
missing or outlier values, i.e. \(x_d^i > \hat{\sigma}_j\) in any
feature and normalize them to have
\(\hat{\mu}_d = 0, \hat{\sigma_d}\). We split the dataset into
\(N_{tr} = 15639\) training and \(N_{test} = 3910\) test examples (80/20
split) and we fit a Neural Network with 3 hidden layers of 256, 128
and 36 units respectively. After 15 epochs using the Adam optimizer
with learning rate \(\eta = 0.02\), the model achieves a normalized
mean squared error (R-Squared) of \(0.25\) (0.75). In depth
information about preprocessing, training and evaluation are provided
in the Appendix.

@@

In real-world datasets, it is infeasible to obtain the ground truth FE
for seamlessly evaluating the competitive methods. For this reason, we
selected the California Dataset, where, in broad-terms, UALE and
PDP-ICE aggree in the quantification of the average effect and of the
uncertainty. In Figure\ref{fig:ex-real-1}, we observe that UALE
complements ALE for the quantification of the heterogeneous effects,
similarly to as ICE complements PDP. Add some content..

\begin{figure}[h]
  \centering
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_1_ale_auto.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_1_pdp_ice.pdf}\\
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_4_ale_auto.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_4_pdp_ice.pdf}\\
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_5_ale_auto.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_5_pdp_ice.pdf}\\
  \caption{Figure 1}
  \label{fig:ex-real-1}
\end{figure}

@@

As in the Syntetic Example (Case 2), we treat as ground-truth the
effects computed on the full training-set \(N=20000\) with dense
fixed-size bin-splitting (\(K=80\)). Given the big number of samples,
we make the hypothesis that the approximation with dense binning is
close to the ground truth. Afterwards, we randomly select \(N=1000\)
samples and we compare UALE approximation with all possible fixed-size
alternatives using \(\mathcal{L}^{\mu}, \mathcal{L}^{\sigma}\). Add discussion

\begin{figure}[h]
  \centering
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_mu_err_feature_1.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_var_err_feature_1.pdf}\\
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_mu_err_feature_4.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_var_err_feature_4.pdf}\\
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_mu_err_feature_5.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_var_err_feature_5.pdf}\\
  \caption{Figure 1}
  \label{fig:ex-real-2}
\end{figure}

\paragraph{Discussion}

\section{CONCLUSION}

>>>>

The California Housing is a largely-studied dataset with approximately
\(20000\) training instances, making it appropriate for robust
Monte-Carlo approximations. The dataset contains \(D=8\) numerical
features with characteristics about the building blocks of California,
e.g. latitude, longitude, population of the block or median age of
houses in the block. The target variable is the median value of the
houses inside the block in dollars that ranges between
\([15, 500] \cdots 10^3\), with a mean value of
\(\mu_Y \approx 201 \cdot 10^3 \) and a standard deviation of
\(\sigma_Y \approx 110 \cdot 10^3\).

We exclude instances with missing values or with outlier values. As
outlier we define the feature values which ar over three standard
deviations away from the mean feature value. We also normalize all
features to zero-mean and unit standard deviation. We split the
dataset into \(N_{tr} = 15639\) training and \(N_{test} = 3910\) test
examples (80/20 split) and we fit a Neural Network with 3 hidden
layers of 256, 128 and 36 units respectively. After 15 epochs using
the Adam optimizer with learning rate \(\eta = 0.02\), the model
achieves a normalized mean squared error (R-Squared) of \(0.25\)
(0.75), which corresponds to a MAE of \(37 \cdot 10^3\) dollars.

Below, we illustrate the feature effect for three features: latitude
\(x_2\), population \(x_6\) and median income \(x_8\). The particular
features cover the main FE cases, e.g.~positive/negative trend and
linear/non-linear curve, and they are appropriate for illustration
purposes. The complete results for all features, along with in-depth
information about the preprocessing, training and evaluation parts are
provided in the Appendix.

@@

In real-world datasets, it is infeasible to obtain the ground truth FE
for seamlessly evaluating the competitive methods. We selected the
particular experiment, because, in broad terms, UALE and PDP-ICE plots
aggree in the estimation of the average effect and
uncertainty. Therefore, we can focus on juding the quality of the
information provided by the two methods.

In Figure\ref{fig:ex-real-1}, we observe, from top to bottom, the
effects for the latitude, population and the median income. The effect
of UALE and PDP-ICE are similar for the population and the median
income. The population has a negative impact that progressively
decreases: from 400 to 1500 people the house value decreases with a
rate of \(-150 (\pm 140)\) dollars per added person, a rate that
decreases from \(-80 (\pm 80)\) to \(-60 (\pm 60)\) dollars per added
person as we move from 1500 to 2800 people. The level of uncertainty
indicates significant variance in absolute value of the rate, but in
the grant majority of instances the rate is negative. With the same
inspection, we observe that the median annual income has a positive
impact on the value (all numbers are thousands of dollars):
\(20\pm 15\) per \(10\) of added median income for incomes in
\([8, 15]\), \(32 \pm 20\) per \(10\) added income in \([15, 60]\) and
\(40 \pm 15\) per \(10\) added income in \([60, 70]\). The uncertainty
indicates that there are less heterogeneous effects about the median
income compared to the number of people. In both cases, we can end-up
to the same conclusion by inspecting the PDP-ICE plots. For the
latitude, there is a small difference in the explanations for the
region \([32, 35]\), where UALE estimates a less negative slope with
less uncertainty than PDP, while the explanations are similar for the
range \([35,39]\), where both methods reveal an increase in the
uncertainty around the feature value \(37.5\).

In general, we observe that UALE complements ALE in the quantification
of the heterogeneous effects, similarly to as ICE complements PDP. The
automatic extraction of constant-effect and constant-uncertainty
regions provided by UALE is helpful for an easier interpretation. On
the other hand, ICE plots sometimes are more descritive on locating
the type of heterogeneity, whereas UALE quantifies only the level (not
the type) of heterogeneity.

\begin{figure}[h]
  \centering
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_1_ale_auto.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_1_pdp_ice.pdf}\\
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_5_ale_auto.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_5_pdp_ice.pdf}\\
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_7_ale_auto.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/feature_7_pdp_ice.pdf}\\
  \caption{Figure 1}
  \label{fig:ex-real-1}
\end{figure}

@@

In this part, we evaluate the robustness of the approximation using
automatic bin-splitting. Following the evaluation framework of
Section~\ref{sec:simulation-examples-2}, we treat as ground-truth the
effects computed on the full training-set \(N=20000\) with dense
fixed-size bin-splitting (\(K=80\)). Given the big number of samples,
we make the hypothesis that the approximation with dense binning is
close to the ground truth. Afterwards, we randomly select less samples
\(N=1000\) and we compare UALE approximation with all possible
fixed-size alternatives, repeating this process for 30 times for
robust results. In Figure~\ref{fig:ex-real-2}, we illustrate the mean
values for \(\mathcal{L}^{\mu}, \mathcal{L}^{\sigma}\) of the 30
repetitions. We observe that automatic bin-spliting provides (close to
the) best approximation in the three features. In the Appendix, we
provide the same evaluation for all features.

\begin{figure}[h]
  \centering
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_mu_err_feature_1.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_var_err_feature_1.pdf}\\
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_mu_err_feature_5.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_var_err_feature_5.pdf}\\
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_mu_err_feature_7.pdf}
  \includegraphics[width=.23\textwidth]{real_dataset_3/compare_var_err_feature_7.pdf}\\
  \caption{Figure 1}
  \label{fig:ex-real-2}
\end{figure}

\section{DISCUSSION}

add limitations.
